Bow shocks, bow waves, and dust waves

### 2018-06-30 papers/Papers1234/sec-dust-bow-intro.tex
Finish description of quasi-parallel cases

As the dusty stream approaches the star, the radiation parameter
\(\Xi\) will increase, with a dependence of \(R^{-2}\) once the stream
is well inside the ionization front.  This increases \(f\rad\)
(eq.~[\ref{eq:frad-Xi}]), but also increases the grain potential,
\(\phi\) (eq.~[\ref{eq:phi-vs-Xi}]) due to the increasing dominance of
grain charging by photoelectric ejection.  Initially, this results in
a lowering of the equilibrium drift velocity to \(w_{10} \sim 0.01\) as
the Coulomb drag kicks in (see Appendix~\ref{sec:cloudy-models-dust}).
However, at smaller radii the slow logarithmic increase in
\(\phi(\Xi)\) means that the drift velocity must start increasing again to
accommodate the linear increase of \(f\rad(\Xi)\).  Eventually,
\(f\rad\) exceeds \(f_{\mathrm{max}}\), the maximum drag force that
proton Coulomb interactions can provide
(eq.~[\ref{eq:fdrag-maximum}]).  This occurs at a critical value of
the radiation parameter, which we denote the \textit{rip point}:
\(\Xi_\dag \sim 1000\).  The variations in \(\Xi_\dag\) with star and grain
parameters, which are of order \SI{+- 0.5}{dex}, are listed in
Table~\ref{tab:Xi-rip} and illustrated graphically in
Figure~\ref{fig:drift-gn}.

@@

In order for a separate outer dust wave to exist, it is necessary for
the grains to decouple from the incoming gas stream before the stream
hits the hydrodynamic bow shock caused by the stellar wind.  The
radius of the rip point, \(R_\dag\), can be expressed in terms of
\(R_*\), the fiducial optically thick bow shock radius introduced in
\S~\ref{sec:strong-gas-grain}:
\begin{equation}
  \label{eq:Rdag-over-Rstar}
  R_\dag = \frac{v_\infty}{\sound}\, \Xi_\dag^{-1/2} R_* \approx v_{10}\, \Xi_\dag^{-1/2} R_* \ ,
\end{equation}
where we have made use of equations~\eqref{eq:Rstar}
and~\eqref{eq:Xi-Prad-over-Pgas}.  The wind bow shock radius is
\(R_0 = \eta^{1/2} R_*\) (eq.~[\ref{eq:x-cases}]), where \(\eta\) is the
wind momentum efficiency (eq.~[\ref{eq:wind-eta-typical}]).
Therefore, the condition \(R_\dag > R_0\) becomes
\begin{equation}
  \label{eq:dust-wave-velocity-condition}
  v_{10} > v_{10,\text{min}} = \bigl( \Xi_\dag \, \eta \bigr)^{1/2} \ . 
\end{equation}
For O stars, the wind efficiency is generally high (\(\eta > 0.1\)) and
\(\Xi_\dag > 2000\) (Tab.~\ref{tab:Xi-rip}), so that dust waves can only
exist when the stream velocity is very high
(\(v_\infty > \SI{150}{km.s^{-1}}\)).  For B~stars, in contrast, the wind
can be much weaker (\(\eta < 0.01\)) and \(\Xi_\dag\) is also smaller, so
that dust waves are permitted by this criterion for much lower stream
velocities: (\(v_\infty > \SI{30}{km.s^{-1}}\)).

@@

The combined effects of the three conditions are illustrated in
Figure~\ref{fig:existence-dust-wave} for each of the three example
main sequence stars from Table~\ref{tab:stars}.  Further restrictions
on the existence of dust waves arise when the effects of magnetic
fields are considered, as will be discussed in
\S~\ref{sec:magn-effects-grain} below.  Note that the three conditions
are restrictions solely on the formation of an \textit{outer} dust
wave, that is, outside of the wind-supported hydrodynamic bow shock.
In the case of the equation~\eqref{eq:dust-wave-low-density-condition}
condition, there is a further possibility: if the gas--grain coupling
(and magnetic coupling) is so weak that it is still unimportant at the
higher densities found in the bow shock shell, then an
inertia-confined \textit{inner} dust wave may form inside the bow
shock, even when \(t_* < \tau_{*,\text{min}}\).  Although the same might
be thought to apply to the condition of
equation~\eqref{eq:dust-wave-velocity-condition}, this is not the
case, since the density compression in the bow shock will reduce the
radiation parameter, \(\Xi\), which moves the rip point, \(R_\dag\), to an
even smaller radius.  Therefore, if radiation has not managed to
decouple a grain before it passes through the shock, it is unlikely to
be able to do it afterwards.

@@

In Figures~\ref{fig:frozen-stream} and~\ref{fig:frozen-trajectories}
we show example results for drag-confined dust waves. Apart from the
inclusion of the magnetic field, the model parameters are the same as
used in Figure~\ref{fig:phase-space-trajectories}, with the exception
that the stream density is increased to \SI{10}{cm^{-3}}.\footnote{The
  reason for using a higher density is to decrease the amplitude of
  the radial oscillations of the trajectories, which allows the dust
  wave structure to be more clearly perceived in the figures. } This
time, we use quasi-parallel (\(\theta\B = \ang{10}\)) and
quasi-perpendicular (\(\theta\B = \ang{75}\)) field orientations.  The
quasi-parallel field is most similar to the non-magnetic case, and the
models shown in the two upper panels of Figure~\ref{fig:frozen-stream}
closely mirror the cases shown in
Figure~\ref{fig:phase-space-trajectories}, with a limit cycle behavior
when \(v_\infty = \SI{40}{km.s^{-1}}\) and stagnant drift when
\(v_\infty = \SI{80}{km.s^{-1}}\).  The principle difference from the 1-D
axial trajectories discussed in \S~\ref{sec:grain-traj-along} is that
the small \ang{10} misalignment of \(\bm{B}\) from the incident stream
direction causes a slow sideways migration, which puts a finite limit
on the time a grain can reside in front of the star.  This can be
appreciated more clearly in Figure~\ref{fig:frozen-trajectories},
which shows the grain position and velocity along a sample streamline
with initial impact parameter \(b = -0.4 R_\dag\) for the two
quasi-parallel models.  In panel~a, we see the same rip-and-snap-back
cycle of de-coupling and re-coupling that was discussed previously,
but only two periods of the cycle are completed before the grain's
lateral migration takes it as far as \(y \approx +R_\dag\), at which point
nothing can stop the gas stream from dragging it past the star and
away.

>>>>

As the dusty stream approaches the star, the radiation parameter
\(\Xi\) will increase, with a dependence of \(R^{-2}\) once the stream
is well inside the ionization front.  This increases \(f\rad\)
(eq.~[\ref{eq:frad-Xi}]), but also increases the grain potential,
\(\phi\) (eq.~[\ref{eq:phi-vs-Xi}]) due to the increasing dominance of
grain charging by photoelectric ejection.  Initially, this results in
a lowering of the equilibrium drift velocity to \(w_{10} \sim 0.01\) as
the Coulomb drag kicks in (see Appendix~\ref{sec:cloudy-models-dust}).
However, at smaller radii the slow logarithmic increase in
\(\phi(\Xi)\) means that the drift velocity must start increasing again to
accommodate the linear increase of \(f\rad(\Xi)\).  Eventually,
\(f\rad\) exceeds \(f_{\mathrm{max}}\), the maximum drag force that
proton Coulomb interactions can provide
(eq.~[\ref{eq:fdrag-maximum}]).  This occurs at a critical value of
the radiation parameter, which we denote the \textit{rip point}:
\(\Xi_\dag \sim 1000\).  The variations in \(\Xi_\dag\) with star and grain
parameters, which are of order \SI{+- 0.5}{dex}, are listed in
Table~\ref{tab:Xi-rip} and illustrated graphically in
Figure~\ref{fig:drift-gn}.  The radius of the rip point, \(R_\dag\), can
be expressed in terms of \(R_*\), the fiducial optically thick bow
shock radius introduced in \S~\ref{sec:strong-gas-grain}:
\begin{equation}
  \label{eq:Rdag-over-Rstar}
  R_\dag = \frac{v_\infty}{\sound}\, \Xi_\dag^{-1/2} R_* \approx v_{10}\, \Xi_\dag^{-1/2} R_* \ ,
\end{equation}
where we have made use of equations~\eqref{eq:Rstar}
and~\eqref{eq:Xi-Prad-over-Pgas}.

@@

In order for a separate outer dust wave to exist, it is necessary for
the grains to decouple from the incoming gas stream before the stream
hits the hydrodynamic bow shock caused by the stellar wind.  The wind
bow shock radius is \(R_0 = \eta^{1/2} R_*\) (eq.~[\ref{eq:x-cases}]),
where \(\eta\) is the wind momentum efficiency
(eq.~[\ref{eq:wind-eta-typical}]).  Therefore, the condition
\(R_\dag > R_0\) becomes from equation~\eqref{eq:Rdag-over-Rstar}:
\begin{equation}
  \label{eq:dust-wave-velocity-condition}
  v_{10} > v_{10,\text{min}} = \bigl( \Xi_\dag \, \eta \bigr)^{1/2} \ . 
\end{equation}
For O stars, the wind efficiency is generally high (\(\eta > 0.1\)) and
\(\Xi_\dag > 2000\) (Tab.~\ref{tab:Xi-rip}), so that dust waves can only
exist when the stream velocity is very high
(\(v_\infty > \SI{150}{km.s^{-1}}\)).  For B~stars, in contrast, the wind
can be much weaker (\(\eta < 0.01\)) and \(\Xi_\dag\) is also smaller, so
that dust waves are permitted by this criterion for much lower stream
velocities: (\(v_\infty > \SI{30}{km.s^{-1}}\)).

@@

The combined effects of the three conditions are illustrated in
Figure~\ref{fig:existence-dust-wave} for each of the three example
main sequence stars from Table~\ref{tab:stars}.  Further restrictions
on the existence of dust waves arise when the effects of magnetic
fields are considered, as will be discussed in
\S~\ref{sec:magn-effects-grain} below.  Note that the three conditions
are restrictions solely on the formation of an \textit{outer} dust
wave, that is, outside of the wind-supported hydrodynamic bow shock.
In the case of the equation~\eqref{eq:dust-wave-low-density-condition}
condition, there is a further possibility: if the gas--grain coupling
(and magnetic coupling) is so weak that it is still unimportant at the
higher densities found in the bow shock shell, then an
inertia-confined \textit{inner} dust wave may form inside the bow
shock, even when \(\tau_* < \tau_{*,\text{min}}\).  Although the same might
be thought to apply to the condition of
equation~\eqref{eq:dust-wave-velocity-condition}, this is not the
case, since the density compression in the bow shock will reduce the
radiation parameter, \(\Xi\), which moves the rip point, \(R_\dag\), to an
even smaller radius.  Therefore, if radiation has not managed to
decouple a grain before it passes through the shock, it is unlikely to
be able to do it afterwards.

@@

In Figures~\ref{fig:frozen-stream} and~\ref{fig:frozen-trajectories}
we show example results for drag-confined dust waves. Apart from the
inclusion of the magnetic field, the model parameters are the same as
used in Figure~\ref{fig:phase-space-trajectories}, with the exception
that the stream density is increased to \SI{10}{cm^{-3}}.\footnote{The
  reason for using a higher density is to decrease the amplitude of
  the radial oscillations of the trajectories, which allows the dust
  wave structure to be more clearly perceived in the figures. } This
time, we use quasi-parallel (\(\theta\B = \ang{10}\)) and
quasi-perpendicular (\(\theta\B = \ang{75}\)) field orientations.  The
quasi-parallel field is most similar to the non-magnetic case, and the
models shown in the two upper panels of Figure~\ref{fig:frozen-stream}
closely mirror the cases shown in
Figure~\ref{fig:phase-space-trajectories}, with a limit cycle behavior
when \(v_\infty = \SI{40}{km.s^{-1}}\) and stagnant drift when
\(v_\infty = \SI{80}{km.s^{-1}}\).

The principle difference from the 1-D axial trajectories discussed in
\S~\ref{sec:grain-traj-along} is that the small \ang{10} misalignment
of \(\bm{B}\) from the incident stream direction causes a slow
sideways migration, which puts a finite limit on the time a grain can
reside in front of the star.  This can be appreciated more clearly in
Figure~\ref{fig:frozen-trajectories}, which shows the grain position
and velocity along a sample streamline with initial impact parameter
\(b = -0.4 R_\dag\) for the two quasi-parallel models.  In panel~a,
corresponding to \(v_\infty = \SI{40}{km.s^{-1}}\), we see the same
rip-and-snap-back cycle of de-coupling and re-coupling that was
discussed previously, but only two periods of the cycle are completed
before the grain's lateral migration takes it as far as
\(y \approx +R_\dag\), at which point nothing can stop the gas stream from
dragging it past the star and away.  All told, the grain remains in
the apex region for about \(1/\sin\theta\B\) times longer than the crossing
time, \(R_\dag / v_\infty\).

In panel~b, corresponding to \(v_\infty = \SI{80}{km.s^{-1}}\), the grain
settles for a while around the stagnant drift radius, \(R_\ddag\), after
the initial rip and turn around.  Again, it slowly migrates sideways,
and eventually recouples to the incident stream, but this time after
reaching \(y \approx +R_\ddag\).  The grain residence time in the apex region,
measured in crossing times, is slightly longer than in panel~a, but is
of the same order.  In both these cases, a second, exterior dense
shell is formed in addition to the hemispherical one produced by the
initial turn around inside \(R_\dag\). In the limit-cycle case, it forms
at the snap-back point, while in the stagnant-drift case, it forms at
\(R_\ddag\) and is significantly denser than the interior shell (upper
right panel of Fig.~\ref{fig:frozen-stream}).
